---
title: "Gene Network Generation Script"
output: html_document
---
Load utility functions and modules
```{r}
library(tidyverse)
library(reshape2)
'%!in%' <- function(x,y)!('%in%'(x,y))
```

*Data Preparation for Use in this Script*

A.For this Funciton a metabolite/Protein/Gene matrix in which:
1. the metabolites/genes/protein are given in the first Column
2. Measurements for each sample as given in following columns
3. All Factors for each sample are separated with an underscore

an example is given below

Symbol		BXD1_HF_Mouse_1			BXD2_HF_Mouse_2			BXD1_HF_Mouse_3
Gene 1		12341					1243					1234
Gene 2		134						134						1234
Gene 3		1243					1243					1243

*IMPORTANT*
the word "Symbol" must be in the cell 
of column 1 row 1

*The Use must define the following*
1. Which of the factors encoded by the underscores is the strain
	(in this case it is strain_factor = 1 because the first factor is strain (BXD1)

2. Can import a dataframe that is formatted as above or a csv
	if a csv use data.type = "csv" and 
	define the file.location = "./path/to/file"

3.Which Gene Network Format is required (either can be uploaded to GeneNetwork without a problem)
	a. format = "row" - the strains are in the first row with the next rows containing the proteins/genes
	b. format = "col" - the strain are given down the first column and the proteins measured in the columns


```{r Gene Network Input, message=FALSE, warning=FALSE}

GeneNetwork_input <- function(data, data.type = "csv", data.location, format = "row",
							  strain_factor = 1, output.file) {

    if (data.type == "csv") {
        inputMatrix <- read.csv(file = data.location, header = TRUE)
    }

    if (data.type == "df") {
        inputMatrix <- read.csv(data , header = TRUE)
    }

inputMatrix.melt <- melt(inputMatrix, id.vars = c("Symbol"))

require(data.table)
inputMatrix.dt <- data.table(inputMatrix.melt, key = "Symbol")
input_dataframe <- as.data.frame(inputMatrix.dt[, head(.SD, 3), by = c("Symbol", "variable")])

input_dataframe[, 1] <- as.character(input_dataframe[, 1])
input_dataframe[, 2] <- as.character(input_dataframe[, 2])
input_dataframe[, 3] <- as.numeric(input_dataframe[, 3])
input_dataframe <- na.omit(input_dataframe)



library(stringr)

n_factor <- str_count(pattern = "_", string = input_dataframe[1, "variable"])

print(paste("There are ", n_factor, " factors encoded with Underscores"))
print("Assuming First Factor is Strain")

    strain.names <- str_split_fixed(input_dataframe$variable, "_", n = 2)[, strain_factor]
input_dataframe$variable <- strain.names

aggres <- input_dataframe %>%
          group_by(Symbol, variable) %>%
          summarize(values = mean(value))

aggres <- aggres %>% filter(variable %!in% "MIX")

colnames(aggres) <- c("Symbol", "variable", "value")

    if (format == "row") {
        gene.network.row <- dcast(aggres, formula = Symbol ~ variable, value.var = "value")
        colnames(gene.network.row) <- gsub(colnames(gene.network.row),
                                   pattern = colnames(gene.network.row[1]),
                                   replacement = "@format=row")

        #Fix Strain names, need to convert . to forward slash 
        #because R converts / into .

        colnames(gene.network.row) <- gsub(colnames(gene.network.row),
                                   pattern = ".", replacement = "/", fixed = TRUE)

        write.table(gene.network.row,
        file = paste(as.character(output.file), "gene.network.row.txt", sep = ""),
		row.names = F, quote = F)
    }


    if (format == "col") {

        gene.network.col <- dcast(aggres, formula = variable ~ Symbol, value.var = "value")

        #Fix Strain names, need to convert . to forward slash 
        #because R converts / into .

        gene.network.col$variable <- gsub(gene.network.col$variable,
                                   pattern = ".", replacement = "/", fixed = TRUE)

		colnames(gene.network.col) <- gsub(colnames(gene.network.col),
                                   pattern = colnames(gene.network.col[1]),
                                   replacement = "@format=column")

        write.table(gene.network.col,
        file = paste(as.character(output.file), "gene.network.row.txt", sep = ""),
		row.names = F, quote = F)

    }


}

```



To generate the file

```{r}
GeneNetwork_input(data.type = "csv",
                  data.location = "2.Trimmed_Row.csv",
                  output.file = "Evan_Row")
 
GeneNetwork_input(data.type = "csv",
                  data.location = "2.Trimmed_Row.csv",
                  output.file = "Evan_Column",
                                                       format = "col")
```



